### 2019-07-04 2-Chapters/3-Chapter/chapter3.tex
A few lines shorter for list of algorithms in 3.3.1

% https://tex.stackexchange.com/a/12430/
\begin{small}
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python}
from SMPyBandits.Arms import *
arm1 = Bernoulli(0.1)
arm2 = Bernoulli(0.5)
arm3 = Bernoulli(0.9)

means = [0.1, 0.5, 0.9]
from SMPyBandits.Environment import MAB
bernoulli_problem = MAB([arm1, arm2, arm3])
# but also
bernoulli_problem = MAB({'arm_type': Bernoulli, 'params': means})

@@

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.75\linewidth]{exampleOfRewards.pdf}
	\caption{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$ (in \textcolor{red}{red}), $0.5$ (in \textcolor{green}{green}) and $0.9$ (in \textcolor{blue}{blue}).}
	\label{fig:3:exampleOfRewards}
\end{figure}

@@

\textbf{Read the Docs.}
Since $2017$, we have been using the free web service provided by Read the Docs (\href{https://readthedocs.org/}{\texttt{ReadTheDocs.org}}).
Read the Docs allows to automatically build the documentation after every commit. This allows to regularly check that the library is well formatted and can be imported correctly, as well as keeping the online documentation up-to-date.
Moreover, they offer to host the documentation online, at \href{https://smpybandits.rtfd.io/}{\texttt{SMPyBandits.ReadTheDocs.io}}.
For more details, see \href{https://readthedocs.org/projects/smpybandits/}{\texttt{ReadTheDocs.org/projects/SMPyBandits}}.
%
Since its first use, the service did about 180 builds, and about $10\%$ of them were useful to detect newly introduced issues or bugs in the code.
The build is configured using the \texttt{.readthedocs.yml} file in SMPyBandits main folder, and it builds the documentation in about two minutes.
I have also been building the documentation manually on a weekly basis, to also host it on \href{https://smpybandits.github.io/}{\texttt{SMPyBandits.GitHub.io}} thanks to GitHub pages.
In the last two years, the documentation saw about 11000 unique visits, while the GitHub project have been visited about 4500 unique visits.

\textbf{Travis CI.}
Since $2018$, we use the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
It allows to automatically run short numerical simulations after every commit,
in order to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments. The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proven very useful to quickly find and fix new bugs.
For more details, see \href{https://travis-ci.org/SMPyBandits/SMPyBandits}{\texttt{Travis-CI.org/SMPyBandits/SMPyBandits}}.
%
Since its first use, the service did about $190$ builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-players model).
Builds typically run for $15$ minutes, and Travis CI have been proven to be very useful in our development process.

@@

    \item
    %  Algorithm 2 is
    Explore-then-Commit \cite{GarivierETC2016} (\texttt{ETC\_KnownGap}),
    % shown in Algorithm~\ref{algo:2:naiveStrategies} above,
    that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems).
    Similarly to $\varepsilon$-greedy, it has a very small footprint in terms of time and memory, but most of the times it only obtains large (linear) regret.

    \item
    %  Algorithm 3 is
    $\mathrm{Exp}3^{++}$ from \cite{Seldin17} (\texttt{Exp3PlusPlus}), using $\alpha=3$ and $\beta=256$ as advised.
    It is a recent anytime variant of the $\mathrm{Exp3}$ algorithm, which was proven to obtain good ``best of both world'' performances.
    % meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.

@@

    \item
    %  Algorithm 5 is
    \klUCB{} from \cite{KLUCBJournal} (\texttt{klUCB}), also shown in Algorithm~\ref{algo:2:indexPolicy} above, using the Bernoulli KL divergence and the corresponding \klUCB{} indexes (coded as \texttt{kullback.klucbBern}).
    % It achieves asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret.
    % with the scaling factor matching Lai and Robbins' lower-bound \cite{LaiRobbins85}.
    It is computationally more costly that \UCB, as at each time step $t$ and for each arm $k$, a convex optimization problem must be solved approximately, in order to compute the $K$ indexes.
    In practice, a few steps of a simple bisection search method are enough to obtain a good numerical precision, and empirically we found that even in the worst cases \klUCB{} is no more than one order of magnitude slower than \UCB.
    % of more than 10 digits
    % \TODOL{Si on en pas parlé avant, ou après dans une annexe, je peux dire un mot ici à propos de l'implémentation en Cython ou en C des fonctions kl et indices kl-UCB !}

@@

In this chapter, we presented our Python library SMPyBandits, by detailing three main points:
$(i)$ its purpose and qualities,
$(ii)$ its organization,
and $(iii)$ an overview of its usage:
%

@@

$(ii)$
We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
As far as now, SMPyBandits is restricted to the finite-arm case, but it supports a wide range of arms distributions.
Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research from December $2016$ to June $2019$.
More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary problem.
All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
%
When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.

@@

\textbf{About the measurement protocol.}
%
To be precise, we used the two following approaches to measure the real time and memory costs of the different algorithms.

@@

\textbf{Comparing measurements.}
%
But what is important in these simulation results is not the exact values, but rather to compare the costs of the different algorithms, and thus only the relative difference in terms of time and memory costs are important.
Relative difference for costs do not depend much on the code and the machine used for the simulation.
%
% The tables~\ref{table:3:time_problem1}, \ref{table:3:time_problem2} \ref{table:3:memory_problem1} and \ref{table:3:memory_problem2} below follow the same structure as the ones given in Section~\ref{sec:3:reviewSPAlgorithms} but report results in terms of computation time (in milli-seconds) and memory (in B).
The computation time is \emph{normalized}, that is it is divided by the horizon, to count the (average) time used by an algorithm at time $t$ from $t=1$ to $t=T$, while the memory cost is \emph{not normalized}.
%
Thus, we can check that efficient algorithms have a running time independent on the horizon $T$, instead of just observing a linear dependency, and we can check that efficient algorithms have a memory cost independent on the horizon.
All the considered algorithms in this Chapter are efficient in this aspect, but in Chapter~\ref{chapter:6} we study algorithms that essentially have to store, and run computations, on an increasing number of observations (\eg, \CUSUMUCB), so their computation cost at time $t$ increases (linearly) with $t$, and thus they are more costly.
%
We can also check that efficient algorithms have a running time and a memory cost linear in the number of arms $K$, while more complex algorithms like BESA suffer from an exponential blow-up of their complexity when $K$ increases.

\textbf{Normalizing data.}
%
The two figures below regret normalized data, in the following way.
For each algorithm, we ran simulations to obtain its (empirical) regret, computation time and space requirement, for different problems with different values of $K$ and $T$.
Simply considering the average of such measurements makes no sense, as for instance two values of the regret for $T=1000$ or $T=50000$ do not have the same order of magnitude.
%
We know that efficient algorithms are expected to follow these patterns:
\begin{itemize}%\tightlist
    \item The final regret $R_T$ should scale as $\cO(K \log(T))$,
    \item The total computation time $\cT_T$ should scale as $\cO(K T)$ (\ie, a computation time of $\cO(K)$ for each time step $t$),
    \item The total memory cost $\cM_T$ should scale as $\cO(K)$, independent of the horizon.
\end{itemize}
Thus, on the one hand, when we show aggregated results from all the different values of $K$ and $T$, we normalized the data by dividing the regrets by $K \log(T)$, the times by $K T$ and the memory by $K$.
%
On the other hand, when we plot a quantity ($R_T$, $\cT_T$, $\cM_T$) as a function of $T$ (resp. of $K$) we only normalize by $K$ (resp. by $T$).

>>>>

% https://tex.stackexchange.com/a/12430/
\begin{small}
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python}
from SMPyBandits.Arms import *
from SMPyBandits.Environment import MAB

arm1 = Bernoulli(0.1) ; arm2 = Bernoulli(0.5) ; arm3 = Bernoulli(0.9)
bernoulli_problem = MAB([arm1, arm2, arm3])

# but also
means = [0.1, 0.5, 0.9]
bernoulli_problem = MAB({'arm_type': Bernoulli, 'params': means})

@@

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.82\linewidth]{exampleOfRewards.pdf}
	\caption{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$ (in \textcolor{red}{red}), $0.5$ (in \textcolor{green}{green}) and $0.9$ (in \textcolor{blue}{blue}).}
	\label{fig:3:exampleOfRewards}
\end{figure}

@@

\paragraph{Read the Docs.}
Since $2017$, we have been using the free web service provided by Read the Docs (\href{https://readthedocs.org/}{\texttt{ReadTheDocs.org}}).
Read the Docs allows to automatically build the documentation after every commit. This allows to regularly check that the library is well formatted and can be imported correctly, as well as keeping the online documentation up-to-date.
Moreover, they offer to host the documentation online, at \href{https://smpybandits.rtfd.io/}{\texttt{SMPyBandits.ReadTheDocs.io}}.
For more details, see \href{https://readthedocs.org/projects/smpybandits/}{\texttt{ReadTheDocs.org/projects/SMPyBandits}}.
%
Since its first use, the service did about 180 builds, and about $10\%$ of them were useful to detect newly introduced issues or bugs in the code.
The build is configured using the \texttt{.readthedocs.yml} file in SMPyBandits main folder, and it builds the documentation in about two minutes.
I have also been building the documentation manually on a weekly basis, to also host it on \href{https://smpybandits.github.io/}{\texttt{SMPyBandits.GitHub.io}} thanks to GitHub pages.
In the last two years, the documentation saw about 11000 unique visits, while the GitHub project have been visited about 4500 unique visits.

\paragraph{Travis CI.}
Since $2018$, we use the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
It allows to automatically run short numerical simulations after every commit,
in order to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments. The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proven very useful to quickly find and fix new bugs.
For more details, see \href{https://travis-ci.org/SMPyBandits/SMPyBandits}{\texttt{Travis-CI.org/SMPyBandits/SMPyBandits}}.
%
Since its first use, the service did about $190$ builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-players model).
Builds typically run for $15$ minutes, and Travis CI have been proven to be very useful in our development process.

@@

    \item
    %  Algorithm 2 is
    Explore-then-Commit (\texttt{ETC\_KnownGap}),
    % shown in Algorithm~\ref{algo:2:naiveStrategies} above,
    that knows the horizon, and a lower-bound on the gap between arms (chosen arbitrarily as $\delta=0.01$, valid for all problems).
    Similarly to $\varepsilon$-greedy, it has a very small footprint in terms of time and memory, but most of the times it only obtains large (linear) regret.

    \item
    %  Algorithm 3 is
    $\mathrm{Exp}3^{++}$ from \cite{Seldin17} (\texttt{Exp3PlusPlus}), using $\alpha=3$ and $\beta=256$ as advised.
    % It is a recent anytime variant of the $\mathrm{Exp3}$ algorithm, which was proven to obtain good ``best of both world'' performances.
    % meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.

@@

    \item
    %  Algorithm 5 is
    \klUCB{} from \cite{KLUCBJournal} (\texttt{klUCB}), also shown in Algorithm~\ref{algo:2:indexPolicy} above, using the Bernoulli KL divergence and the corresponding \klUCB{} indexes (coded as \texttt{kullback.klucbBern}).
    % It achieves asymptotical optimal problem-dependent bounds with a $\cO(\log(T))$ regret.
    % with the scaling factor matching Lai and Robbins' lower-bound \cite{LaiRobbins85}.
    It is computationally more costly that \UCB,
    % as at each time step $t$ and for each arm $k$, a convex optimization problem must be solved approximately, in order to compute the $K$ indexes.
    % In practice, a few steps of a simple bisection search method are enough to obtain a good numerical precision,
    even if empirically we found that even in the worst cases \klUCB{} is no more than one order of magnitude slower than \UCB.
    % of more than 10 digits
    % \TODOL{Si on en pas parlé avant, ou après dans une annexe, je peux dire un mot ici à propos de l'implémentation en Cython ou en C des fonctions kl et indices kl-UCB !}

@@

We chose to present here our Python library SMPyBandits in this chapter rather than in the Appendix, because all the numerical experiments on single-player multi-arm bandits in Chapter~\ref{chapter:25}, and on more sophisticated models in the last two Chapters~\ref{chapter:5} and \ref{chapter:6} are built on our library.
%
We detailed three main points of SMPyBandits:
$(i)$ its purpose and qualities,
$(ii)$ its organization,
and $(iii)$ an overview of its usage:
%

@@

$(ii)$
We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
As far as now, SMPyBandits is restricted to the finite-arm case, but it supports a wide range of arms distributions.
Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research from December $2016$ to June $2019$.
More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary problem.
All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
%
When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate, mean cumulated reward, real time or memory costs.

@@

\paragraph{About the measurement protocol.}
%
To be precise, we used the two following approaches to measure the real time and memory costs of the different algorithms.

@@

\paragraph{Comparing measurements.}
%
But what is important in these simulation results is not the exact values, but rather to compare the costs of the different algorithms, and thus only the relative difference in terms of time and memory costs are important.
Relative difference for costs do not depend much on the code and the machine used for the simulation.
%
% The tables~\ref{table:3:time_problem1}, \ref{table:3:time_problem2} \ref{table:3:memory_problem1} and \ref{table:3:memory_problem2} below follow the same structure as the ones given in Section~\ref{sec:3:reviewSPAlgorithms} but report results in terms of computation time (in milli-seconds) and memory (in B).
The computation time is \emph{normalized}, that is it is divided by the horizon, to count the (average) time used by an algorithm at time $t$ from $t=1$ to $t=T$, while the memory cost is \emph{not normalized}.
%
Thus, we can check that efficient algorithms have a running time independent on the horizon $T$, instead of just observing a linear dependency, and we can check that efficient algorithms have a memory cost independent on the horizon.
All the considered algorithms in this Chapter are efficient in this aspect, but in Chapter~\ref{chapter:6} we study algorithms that essentially have to store, and run computations, on an increasing number of observations (\eg, \CUSUMUCB), so their computation cost at time $t$ increases (linearly) with $t$, and thus they are more costly.
%
We can also check that efficient algorithms have a running time and a memory cost linear in the number of arms $K$, while more complex algorithms like BESA suffer from an exponential blow-up of their complexity when $K$ increases.

\paragraph{Normalizing data.}
%
The two figures below regret normalized data, in the following way.
For each algorithm, we ran simulations to obtain its (empirical) regret, computation time and space requirement, for different problems with different values of $K$ and $T$.
Simply considering the average of such measurements makes no sense, as for instance two values of the regret for $T=1000$ or $T=50000$ do not have the same order of magnitude.
%
We know that efficient algorithms are expected to follow these patterns:
\begin{itemize}%\tightlist
    \item The final regret $R_T$ should scale as $\cO(K \log(T))$,
    \item The total computation time $\cT_T$ should scale as $\cO(K T)$ (\ie, a computation time of $\cO(K)$ for each time step $t$),
    \item The total memory cost $\cM_T$ should scale as $\cO(K)$, independent of the horizon.
\end{itemize}
Thus, on the one hand, when we show aggregated results from all the different values of $K$ and $T$, we normalized the data by dividing the regrets by $K \log(T)$, the times by $K T$ and the memory by $K$.
%
On the other hand, when we plot a quantity ($R_T$, $\cT_T$, $\cM_T$) as a function of $T$ (resp. of $K$) we only normalize by $K$ (resp. by $T$).

### 2019-07-04 2-Chapters/Conclusion/conclusion.tex
A few lines shorter for list of algorithms in 3.3.1

As detailed in Chapter~\ref{chapter:2}, we focused in this thesis on MAB problems with a finite number of arms, and binary, bounded or real-valued rewards.
Thus two possible directions of future research could be to extend our works
to MAB models with either multidimensional rewards, like contextual bandits, or infinite arms, like Lipschitz bandits.
The hypotheses we made were motivated by the target applications, but these other models have also started to be used for cognitive radio applications.

>>>>

As detailed in Chapter~\ref{chapter:2}, we focused in this thesis on MAB problems with a finite number of arms, and binary, bounded or real-valued rewards.
Thus two possible directions of future research could be to extend our works
to MAB models with either multidimensional rewards, like contextual bandits \cite{Li10,Luo18,ChenLeeLuoWei2019}, or infinite arms \cite{valko2016bandits,Combes17}, like Lipschitz bandits.
The hypotheses we made were motivated by the target applications, but these other models have also started to be used for cognitive radio applications.
